---
title: "Results"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    df_print: kable
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup,  message=FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, warning = FALSE, fig.width = 16/2, fig.height = 9/2
)
```

## Results

```{r} 
#starter code :)
#loading libraries
library(tidytext)
library(stringr)
library(tidyverse)
library(wordcloud)
library(dplyr)
library(reshape2)
library(lubridate)
library(ggplot2)

#load the rda file
load(file = "textdata.Rda")

#code to get started
#"pasted this part from Emily :), original code by Jeny :)"
df_tidy <- df_full %>% 
  unnest_tokens(word, content) %>% 
  anti_join(stop_words) 
```


## Negative Sentiment of 2020 Emails 

```{r}
#calls for bing sentiment lexicon 
bing_lex = get_sentiments("bing")

#each year splits into groups of 50 words (we can vary this)
#sentiment score is calculated
bing_score <- df_tidy %>%
  mutate(year = year(date)) %>%
  group_by(year) %>% 
  mutate(word_count = 1:n(),
         index = word_count %/% 50 + 1) %>% 
  inner_join(bing_lex, by = "word") %>%
  count(year, index = index , sentiment) %>%
  ungroup() %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative,
         date = factor(year)) %>%
  ggplot(aes(index, sentiment, fill = year)) +
  geom_bar(alpha = 0.5, stat = "identity", show.legend = FALSE) +
  facet_wrap(~ date, ncol = 2, scales = "free_x")
bing_score
```

We were interested to see how the sentiment score of the grouped content, consisting of $50$ words each, changes as time passes within the four years. For the most part, emails communicated in $2020$ were positive overall, however, over time the emails begin to have content that the bing lexicon determines as negative. In our raw data, COVID-19 related emails began February $7$th, $2020$, and since then the sentiments of the emails have been negative. There have also been more words contained in the $2020$ emails in comparison to the other three years. The only other year that has a similar negative sentiment as $2020$ is $2018$, and a significant number of $2018$ emails dealt with campus climate and issues of inclusion and equity. Clearly, according to the bing lexicon, communication between the administration and the Smith community is more likely to be negative than in the other three years.

## Visualization 2


## Visualization 3